Ecosystem carbon balance in the Hawaiian Islands under different scenarios of future climate and land use change

### 2021-08-12 Manuscript/journal_correspondence/ERL_responseletter.Rmd

\blankline

Dear Editors:

\blankline

Please consider this revised version of article ERL-111271, "Ecosystem carbon balance in the Hawaiian Islands under different scenarios of climate and land use change", for publication in _Environmental Research Letters_. We thank the two referees for their review of this manuscript, and have fully addressed each of their comments in detail below. Our point-by-point responses immediately follow each referee comment, and we include line numbers of the revised manuscript in each response to identify where we made changes to the text.

@@

\raggedright
Paul C. Selmants, Ph.D. 
\linebreak
Research Ecologist 
\linebreak
U.S. Geological Survey

## Referee 1, comments to the author:

**The manuscript describes an impressive effort projecting ecosystem carbon balance in the Hawaiian Islands under climate change and land use change scenarios. This assessment of future regional terrestrial carbon dynamics can provide a support to local planning of climate mitigation strategies through ecosystem carbon sequestration. The manuscript is well written and organized. However, more clarification and justification are needed for methodology and assumptions about how future NPP and land cover change is modelled. And also, more details are needed such as maps illustrating how projected change in carbon flux (e.g., NPP and NBP) and carbon pools (live biomass, soil organic soil) is spatially distributed. I have provided detailed comments below.**

As suggested by this reviewer, we have added clarification and justification about how future NPP and land cover change were modeled, as well as new figures illustrating the spatial distribution of current and projected future carbon fluxes. Our responses to individual comments below indicate where we made these changes.

**L109, why need to assign each cell a state type? What is consideration for it?**  The state of each cell tracks a range of information about that cell, including land cover class, moisture zone, island, age, and time since transition. This set of information is used to define the probabilities that each cell can undergo a particular type of transition, as well as how often each of these transitions can occur. We have added text and references to the first paragraph in the Methods section that defines the state of each cell (lines xxx).

**L114, transitions between ‘state type’? Is it ‘land cover classes’? According to the definition of state type as unique combination of moisture zones, islands and land cover classes at L110, is there transition from dry forest and mesic forest in the same island?**  
We agree this was unclear. We have added text to the first paragraph of the Methods section that defines "transition" (line xxx) and we have re-written this paragraph to make clear that transition pathways define which land cover classes can be converted to other land cover classes (lines xxx).

**L123-126, high land use scenario samples historical rate between median and maximum, but why not low land use scenario sample rates between minimum and median? What is the justification that low land use scenario will have rate lower than historical records?**  
Sampling from this range of agricultural contraction and expansion rates in the low land use scenario shifts the balance toward agricultural contraction, thus reducing over time the total land area under active cultivation in the Hawaiian Islands. We consider this to be a reasonable possibility under a low land use scenario, especially given the volatility in agricultural land area change experienced in the Hawaiian Islands over the recent historical period (1992-2010). We have added text that justifies bounding agricultural land area change rates between zero and the minimum 1992-2010 historical rates in the low land use scenario (lines xxx - xxx).

**L126, missing description for urbanization rate for high land use scenario.**  
Transition rates for urbanization in the high land use scenario were sampled from a uniform distribution bounded by the median and maximum historical rate, as stated on lines 121-123 of the original manuscript. We have re-worded this description to make it more clear that it includes urbanization (lines xxx - xxx).

**L133, what transitions can increase forest area such as low land use scenario shown in figure S5? Does tree plantation refer to reforestation or afforestation? If afforestation, what is the source land cover class for this plantation?** 
Forest area increases in the low land use scenario because of agricultural contraction (i.e., abandonment of previously cultivated land). We have added text to the supplementary explaining that the increase in forest, grassland, and shrubland area in the low land use scenario between 2020 and 2040 is largely due to agricultural contraction over the same time period (supplementary lines xxx - xxx). Tree plantation does not refer to reforestation or afforestation. We did not include expansion of tree plantations as a transition type, and so there is no land cover class that can be converted to tree plantation in either land use scenario. Once harvested, tree plantations can either be replanted to tree plantations or be converted to grassland or agriculture, which partially explains the increase in grassland area in both land use scenarios (supplemental Figure 6). We have added text to the Methods section that better describes tree plantation harvest rates and transitions (lines xxx-xxx). The source of all land cover classes, including tree plantations, is the land cover class map depicted in Figure 1, derived from Jacobi et al. (2017).

@@

**L136, it will be helpful to add average rotation cycle corresponding to the 75% and 40% historical harvest rate? Like how soon planted trees will be harvested in low and high land use scenarios.**  
We have added text defining the historical harvest rotation rate as between 5-7 years, as well as more detail on how tree plantation area is projected to change under both land use scenarios (lines xxx - xxx).

**L137, will wildfire take place only in forests? If so, what is the resulting land cover class? Would it still recover back to forest after a certain time?**  
Wildland fire in Hawaiʻi primarily ignites in non-native grasslands and shrublands, but can sometimes spread into forests. We have added text and references that describe where wildland fire occurs in the Hawaiian Islands (lines xxx - xxx).

**L140, is state type land cover classes? Since ‘state type’ is defined as combination of moisture zone, island, and land cover type at line 110, how can ‘state type’ be re-combined with moisture zone and island here?**  
We agree this was a confusing use of terminology. We have changed "state type" to "land cover class" here (line xxx) and elsewhere throughout the manuscript when it refers only to land cover classification. We now use the term "state type" only when referring to the unique combination of island, moisture zone, and land cover class.

**L158, moisture zone is included in state type as defined at line L110, isn’t it?**  
This was a confusing use of terminology. We have omitted "moisture zone" here and now only refer to "state type" in this sentence (line xxx) to be consistent with our previously stated definition of "state type" as the unique combination of island, moisture zone, and land cover class (line xxx).

**L162, how litter and dead wood pools were initialized?**  
We have added text describing how litter and dead wood pools were initialized using values from the IBIS DGVM (lines xxx-xxx).

**L161, what is 30-year climate normal come from? Does it have consistent temperature and rainfall distribution with RCP projections during 2010 and 2020? Or is there any bias between 30-year climate normal and RCP projections?**  
The contemporary 30-year climate normals for rainfall and temperature come from the references cited (Giambelluca et al. 2013, 2014), as depicted in supplementary Figure 3. We have added text that better describes statistical downscaling of CMIP5 temperature and rainfall projections and how they compare to contemporary climate data (supplementary lines xxx-xxx).

**L163, figure 2 shows total live biomass in 2010 is about 68 Tg C, but total forest aboveground biomass is estimated as 36 Tg C in Asner et al 2016 which is used for initialization here. Where is the additional 32 Tg C from? What is ratio of aboveground biomass to total live biomass in IBIS?**  
Asner et al. (2016) only quantifies aboveground forest live biomass. The additional 32 Tg C comes from the sum of belowground forest live biomass and total live biomass in grasslands, shrublands, and agriculture. We have added text and citations describing how we used a power function applied to the Asner et al. (2016) abovegrould forest live biomass values to estimate forest belowground live biomass and, subsequently, estimate total forest live biomass, which was used to calibrate the IBIS DGVM (lines xxx-xxx).

**L163, what is calibrated GPP NPP rate? It will be helpful to list out the mean or the range.**
We have added text (lines xxx-xxx) listing the range of GPP values from the calibration dataset (Kimball et al. 2017) and added a supplementary table of IBIS-derived mean NPP rates for each unique combination of moisture zone and land cover class (supplementary Table 1).

**L166-168, is similar multiplier also applied to living biomass? If not, will this single NPP multiplier break the equilibrium between NPP and biomass? For example, for a simulation cell, NPP and biomass should reach dynamic equilibrium in IBIS simulation, a given NPP could only support carbon turnover of a certain number of trees. External decease NPP by a multiplier will in turn reduce biomass in order to match it, causing artificial reduction in biomass.**  
No, there is no set of spatial multipliers applied to living biomass. We have added text and figures to the supplementary that better describes how we derived and applied the NPP multipliers (supplementary lines xxx - xxx, supplementary figures 5 & 6). We have also added text to the manuscript describing how the LUCAS initialization of carbon stocks and fluxes allowed for these stocks and fluxes to reach a dynamic equilibrium across the landscape of the Hawaiian Islands (lines xxx - xxx).

**Line 173, why assume Rh as ratio of NPP? Is the ratio constant over time and across climate change scenarios? Why not to estimate Rh based on soil moisture, temperature and soil carbon stocks? Because of lack of soil temperature in RCP 4.5 and 8.5? Need more justification for this flux ratio method.** 
We have added text and citations justifying the fact that plant productivity is the primary factor driving variation in ecosystem carbon loss at annual timescales and that the vast majority of annual carbon loss from soils was fixed by photosynthesis within that year (lines xxx - xxx). Although we agree that gridded annual maps of projected future soil moisture and soil temperature could potentially improve predictions of ecosystem carbon flux in our model, these data products do not exist for the Hawaiian Islands.

**L177-179, are the growth and decay multipliers varied at annual basis? How are multipliers derived for time periods CMIP5 projections do not cover? Such as period between 2020-2049? It is very necessary to include the distribution of these multipliers, are they enhancing or weakening NPP over time at two RCPs?** 
Yes, the growth and decay multipliers vary annually. We have added text, citations, and figures that better describes how growth and decay multipliers were derived for intermediate time periods and how they vary on an annual basis (lines xxx-xxx, supplementary lines xxx-xxx, supplementary Figures 5 & 6). The spatial distribution for the contemporary NPP spatial multipliers is now depicted in supplementary Figure 5, and the spatial distribution of projected future NPP multipliers under both RCP 4.5 and RCP 8.5 are now depicted in supplementary Figure 6.

**L182, the empirical model used here to relate NPP to temperature and rainfall is core to projections of carbon balance. It is necessary to present this empirical relationship in the paper, such as what the optimal temperature of NPP is and how it changes with rainfall. For example, figure S4 shows projections of temperature and rain, there should be corresponding projections of NPP change.**  
We have added a supplemental figure displaying the spatial distribution of projected future NPP multipliers based on future climate projections (supplemental Figure 6). We have also added text to the supplementary that better describes how we employ the empirical NPP equations to calculate spatial NPP anomalies for both contemporary and projected future NPP (supplemental lines xxx).

**L184, what is the temporal coverage and resolution of these climate projections? Do them only cover the period of mid-century (2049-2069) and end-of-century (2070-2099) as described in L179?**  
Yes, climate projections only cover the period of mid-century (2049-2069) and end-of-century (2070-2099) as described on line 179 of the original manuscript. We have added text to the supplementary Climate section that provides more detail on climate projections (supplementary lines xxx).

**L239: is the NPP decrease during 2010-2020 evident in satellite observations? To what extent of this decrease is caused by climate conditions change rather than by forest area changes?**  
Statewide NPP declined by approximately 2% on average between 2010 and 2020 across all scenarios and Monte Carlo iterations (8.13 Tg y^-1^ in 2010 to 7.95 Tg y^-1^ in 2020), and the ranges of statewide annual NPP estimates broadly overlapped for each of these years (2010 range = 7.96 - 8.28 Tg ^y-1^, 2020 range = 7.85 - 8.1 Tg ^y-1^). This slight decline in modeled NPP over 10 years, which is well within the error range of coarse resolution satellite observations, was due almost entirely to land cover change - which includes but is not limited to forest area change. We have added text that describes this change in NPP over the contemporary (2010-2020) time period to the discussion section (lines xxx-xxx).

@@

**L242. It is misleading to state ‘climate change led to increased Rh over time’ as figure 3 shows decreasing trend of Rh under high land use scenario and two RCP climate change scenarios. It should be revised as something like ‘switching from RCP 4.5 to 8.5 led to increased Rh’.**
We agree, this phrasing was misleading. We have altered the text as suggested by this referee (lines xxx).

**L244, why the increase in total soil carbon stock (figure 2) will not enhance Rh? Is it due to the assumption made by this work that Rh is only ratio of NPP and latter decrease over time? What are drivers of Rh decrease?**  
Rh only declines consistently under the high land use scenario in both RCPs (Figure 3). In the low land use scenario Rh remains relatively stable after 2060 under RCP 4.5 and increases after 2020 under RCP 8.5. As stated previously, we have added text and citations justifying NPP as a direct driver of R~h~ because the vast majority of carbon respired heterotrophically from terrestrial ecosystems was fixed by photosynthesis that same year (see lines xxx of revised manuscript). Also, soil carbon stocks do not increase consistently under all scenarios. Under RCP 8.5, soil carbon stocks begin to asymptote in the latter half of the 21st century in the low land use scenario and begin to decline slightly in the high land use scenario. The drivers of R~h~ change over time are described on lines 241-245 of the original manuscript (lines xxx of the revised manuscript).

**L244, why Rh shows very sharp decrease before 2020 under all land use scenarios with all RCPs, such decrease rate seems to be higher than the rest period. Are these decreases supported by observations?**  
Land use change rates are the same in all four scenarios from 2010-2020 and climate driven changes in R~h~ are still small during this initial time period, which explains why R~h~ behaves consistently across these four scenarios. We have added text to make that clear in the Methods section (lines xxx). The decrease in R~h~ before 2020 is driven almost entirely by land use change, specifically the increase in urban area at the expense of natural and agricultural land (supplementary Figure 7). We have added text that describes these trends (supplementary lines xxx).

**Line 341, what is the source of ‘Hawai’i-specific climate data’?**
The Hawaiʻi specific climate data is from Giambelluca et al. (2013), Giambelluca (2014), Elison Timm et al. (2015), and Elison Timm et al. (2017) - all of which are cited several times throughout the manuscript and are depicted in supplementary figures 3 and 4. We have added citations to these data again here in response to this comment (lines xxx).

@@

## Referee 2, comments to the author:

**The manuscript entitled “Ecosystem carbon balance in the Hawaiian Islands under different scenarios of future climate and land use change” is interesting and suitable for publication but several mistakes in the manuscript should be corrected before acceptance.**

**Specific comments**

**Authors should carefully recheck the manuscript for the typos and also write the meaning of all acronyms because there are some which are not presented.**  
We have carefully checked the manuscript for typographical errors and changed "our" to "Our" on line 88, changed "600pp" to "600 ppm" on line 201, and corrected the spelling of "supplemental" on lines 61, 157, 240, and 241 of the revised manuscript. We have also confirmed that all acronyms are spelled out upon initial use in the main body of the manuscript.

**All the figures are clearly presented but authors must check all the figure pixels again. Some of them are not very clear.**  
We have increased the resolution of all figures to 400 dpi in response to this comment.

@@

**I suggest that authors check the bibliographic references in the main text and revise the format of documents to meet the requirement of the Journal.**
We are using the Institute of Physics (IOP) citation style language (csl), which meets the guidelines for this Journal. No changes were made in response to this comment.

**In the abstract, please add more information about the methods of the experiment.**  We have added text to the abstract that provides more information about the methods of this project (lines xxx).

**In the abstract, please don't write the same conclusion lines as they were written in the main conclusion section.**  
We have altered the text so that the conclusions in the abstract are not written identically to those in the Conclusions section (lines xxx).

**Hypothesis should be given. How this work is different from the available literature?**
A hypothesis would be warranted if the purpose of this study was to test existing theory using a controlled experiment, which was not the case. The purpose of this study was to use existing theory to make projections about the future, and so we believe a statement of objectives is sufficient. We describe how this work compares to the available literature in the Discussion section (e.g., lines xxx - xxx), as recommended by the author guidelines for this Journal. No changes were made in response to this comment.

**Objectives should be rewritten in detail and comprehensively.** 
We believe we have stated the objectives of this study clearly and succinctly, both in the abstract and at the end of the Introduction section. That said, we have altered the text to add more detail to our stated objectives (lines xxx-xxx).

**How did the authors perform the calibration and validation process?**
We have added text describing the calibration of the model in the Methods section on lines xxx-xxx. Validation of ecosystem carbon balance estimates in the Hawaiian Islands is problematic given the paucity of data, as we describe in detail the Discussion section, however we were able to compare our results to those from previous studies (lines xxx-xxx).  
**How did they select the sensitive parameters of the model that they used?**
It is unclear to us what the referee is referrring to in this question, as they do not point to a particular portion of the manuscript. We assume they are referring to our test of model sensitivity to different rates of a CO~2~ fertilization effect (CFE). We describe how parameters were selected for this CFE sensitivity test in the Methods section on lines xxx-xxx of the revised manuscript. Without further clarification, we do not know how to better respond to this question. No changes were made in response to this comment.

**Discussion, the discussion part is not related to the results part. Why you discuss these? What does it have to do with the result?** 
We followed the author guidelines for this journal by discussing the significance of the results and comparing them with previous work using relevant references, and so we disagree with this referee that our discussion section is unrelated to the results section.

### 2021-08-12 Manuscript/HI_Model_ms.Rmd
updated manuscript and supplementary files

## Abstract
The State of Hawaiʻi passed legislation to be carbon neutral by 2045, a goal that will partly depend on carbon sequestration by terrestrial ecosystems. However, there is considerable uncertainty surrounding the future direction and magnitude of the land carbon sink in the Hawaiian Islands. We used the Land Use and Carbon Scenario Simulator (LUCAS), a spatially explicit stochastic  simulation model that integrates landscape change and carbon gain-loss, to assess how projected future changes in climate and land use will influence ecosystem carbon balance in the Hawaiian Islands under all combinations of two radiative forcing scenarios (RCPs 4.5 and 8.5) and two land use scenarios (low and high) over a 90-year timespan from 2010-2100. We used simulation modeling to assess how projected future changes in climate and land use will influence ecosystem carbon balance in the Hawaiian Islands under all combinations of two radiative forcing scenarios (RCP 4.5 and RCP 8.5) and two land use scenarios (low and high) over a 90-year timespan from 2010-2100. Collectively, terrestrial ecosystems of the Hawaiian Islands acted as a net carbon sink under low radiative forcing (RCP 4.5) for the entire 90-year simulation period, with low land use change further enhancing carbon sink strength. In contrast, Hawaiian terrestrial ecosystems transitioned from a net sink to a net source of CO~2~ to the atmosphere under high radiative forcing (RCP 8.5), with high land use accelerating this transition and exacerbating net carbon loss. A sensitivity test of the CO~2~ fertilization effect on plant productivity revealed it to be a major source of uncertainty in projections of ecosystem carbon balance. Greater mechanistic understanding of plant productivity responses to rising atmospheric CO~2~ will be essential to realistically constrain simulation models used to evaluate the effectiveness of ecosystem-based climate mitigation strategies.

@@

The State of Hawaiʻi exemplifies the challenges associated with projecting the interactive effects of future climate and land use change on ecosystem carbon balance at a regional scale. Hawaiʻi was the first U.S. state to enact legislation committing to full carbon neutrality, requiring the state to account for and offset all of its greenhouse gas emissions by 2045 (State of Hawaiʻi Acts 15 and 16). This legislation emphasizes the mitigation potential of natural ecosystems as a key component to emissions reduction, necessitating baseline estimates and future projections of land carbon sink strength. However, Hawaiʻi's challenging terrain complicates these assessment efforts. The main Hawaiian Islands are a complex mosaic of natural and human-dominated landscapes overlain by steep climate gradients across relatively short distances, with mean annual temperature ranging from ~4-24° C [@giambelluca_evapotranspiration_2014] and mean annual rainfall ranging from ~200-10,200 mm [@giambelluca_online_2013]. Temperatures have risen rapidly in the Hawaiian Islands since the mid 1970s [@giambelluca_secular_2008; @mckenzie_temperature_2019] and a long-term drying trend has persisted since the early 1920s [@frazier_spatial_2017], resulting in reduced forest biomass and productivity [@barbosa_effects_2017]. These same drying and warming trends have increased the frequency and intensity of wildland fire [@trauernicht_contemporary_2015; @trauernicht_vegetationrainfall_2019] with predictable negative effects on ecosystem carbon balance [@selmants_baseline_2017]. Ecosystem carbon stocks across the main Hawaiian Islands have also been strongly influenced by the legacy of past land use change [@osher_effect_2003; @asner_high-resolution_2011]. Thousands of hectares were deforested beginning in the late 19th century to clear land for sugar plantations and cattle pasture [@cuddihy_alteration_1990]. Since the mid-20th century, much of this agricultural land has been steadily converted to urban areas, commercial forestry plantations, or simply abandoned and colonized by non-native grass species [@suryanata_diversified_2009; @perroy_evolving_2016]. Although these past trends surely inform the future impact of climate and land use change on ecosystem carbon balance, high spatial and temporal heterogeneity complicates realistic projection efforts. To date only one study has attempted to integrate land use, climate, and natural disturbances into future projections of Hawaiian ecosystem carbon balance, with projections limited to the mid-21st century under a single land use change scenario and moderate radiative forcing [SRES A1B, equivalent to RCP 6; @selmants_baseline_2017].

We used a stochastic, spatially explicit simulation model to estimate ecosystem carbon balance for Hawaiʻi's natural and agricultural lands on an annual basis for the period 2010–2100 under a range of assumptions about future climate, land use, land cover, disturbance, and global CO~2~ emissions [@daniel_state-and-transition_2016; @daniel_integrating_2018; @sleeter_effects_2019]. We developed four unique scenarios that explored different pathways, or future trajectories, of land use change and climate change. These four scenarios represent all combinations of two land use change pathways (low and high) and two radiative forcing pathways (representative concentration pathway [RCP] 4.5 and RCP 8.5). In addition to these four scenarios, we conducted a separate series of simulations to examine how ecosystem carbon balance estimates vary in response to different levels of a CO~2~ fertilization effect (CFE) on net primary productivity [NPP; @sleeter_effects_2019]. Our goals were to estimate changes in Hawaiian ecosystem carbon balance and their uncertainties under a range of plausible future scenarios, quantify the relative impact of major controlling processes such as land use change, disturbance, and climate change, and assess the sensitivity of model estimates to the introduction of a CFE on NPP.

@@

### *2.1 Study area*
The spatial extent of this study was the terrestrial portion of the seven main Hawaiian Islands (figure 1), a total land area of 16,554 km^2^. We subdivided this landscape into a grid of 264,870 simulation cells, each of which was 250 x 250 m in size. Each simulation cell was assigned to one of 210 possible state types based on the unique combination of three moisture zones (dry, mesic, and wet; supplemental figure 1), seven islands, and ten discrete land cover classes (figure 1). We also tracked the age and time since transition for each simulation cell as continuous state variables [@daniel_state-and-transition_2016].

```{r fig.align = 'center', out.width = "99%", fig.cap = "Land cover classification of the seven main Hawaiian Islands, adapted from Jacobi \\textit{et al.} (2017). Agriculture in this map combines herbaceous and woody crops, but these two crop types are treated as separate land cover classes in the simulation model. Water and Wetland land cover classes are not shown."}
knitr::include_graphics("./fig_images/fig1_HI_LC.png")
```

### *2.2 States and transitions*
We developed two land use scenarios (low and high) with transition pathways modified from @daniel_state-and-transition_2016. Transition pathways were pre-defined to represent urbanization, agricultural contraction, agricultural expansion, harvesting of tree plantations, and wildfire. Agriculture, forest, grassland, shrubland, and tree plantation land cover classes each had multiple transition pathways, i.e., they could each be converted into a variety of other land cover classes. The barren land cover class could only transition to developed, so its sole transition pathway was urbanization. Although most land cover classes had an urbanization transition pathway, there was no transition pathway out of an urbanized (developed) state. Water and wetland land cover classes remained static throughout the simulation period.

Transition targets set the area to be transitioned over time. These targets were based on recent historical trends of land use change in the Hawaiian Islands from 1992-2011 [@noaa_coastal_2020] and on population projections for the State of Hawaiʻi [@kim_population_2018]. For the high land use scenario, transition rates for each timestep and Monte Carlo realization were sampled from uniform distributions bounded by the median and maximum recent historical rates of agricultural contraction, agricultural expansion, and urbanization for each island. For the low land use scenario, rates of agricultural contraction and expansion were sampled from uniform distributions bounded by zero and the minimum recent historical rates for each island. Bounding agricultural change rates between zero and the recent historical minimum shifts the balance toward agricultural contraction over time in the low land use scenario, which reduces the land area under active cultivation and contrasts with the high variability in agricultural land change rates during the recent historical past [@suryanata_diversified_2009]. Urbanization rates in the low land use scenario were based on island-level population estimates and projections at five year intervals from 2010-2045 [@kim_population_2018]. We converted population projections into urbanization transition targets following @sleeter_future_2017 by calculating population density for each island and then projecting future developed area based on the five-year incremental change in island population. The spatial extent of agricultural contraction, agricultural expansion, and urbanization was constrained in both land use scenarios based on existing zoning maps [@daniel_state-and-transition_2016]. Transition targets for tree plantation harvest were set at ~75% of recent historical rates in the high land use scenario and ~40% of recent historical rates in the low land use scenario [@daniel_state-and-transition_2016]. Tree plantation forestry in the State of Hawaiʻi is primarily short-rotation (5-7 years) *Eucalyptus* spp., harvested at a rate of approximately 2 km^2^ y^-1^ over the past decade. In both the high and low land use scenarios, approximately 60% of tree plantation harvests were replacement harvests resulting in conversion to grassland or agriculture. The remaining 40% were rotation harvests replanted to *Eucalyptus* spp. There was no transition pathway from any land cover class into tree plantation in either land use scenario, which is consistent with recent historical trends of stable or declining tree plantation land area.

@@

Initial carbon stocks and baseline carbon flows were derived from the Integrated Biosphere Simulator [IBIS; @foley_integrated_1996; @liu_critical_2020], a process-based dynamic global vegetation model. We initiated IBIS with minimal vegetation and simulated forward for 110 years using 30-year climate normals for the Hawaiian Islands [supplemental Figure 3; @giambelluca_online_2013; @giambelluca_evapotranspiration_2014]. We calibrated IBIS carbon stocks with statewide gridded datasets of soil organic carbon [@ssurgo_2016] and forest aboveground live biomass [@asner_rapid_2016] converted to total forest live biomass by applying a power function to calculate the belowground portion [@mokany_critical_2006] as described in @selmants_baseline_2017. We calibrated gross photosynthesis in IBIS using a Hawaiʻi-specific gridded dataset of gross primary production (GPP) derived from MODIS satellite imagery [@kimball_evaluating_2017]. Mean annual GPP from @kimball_evaluating_2017 ranged from 0.123 - 3.88 kg C m^-2^ y^-1^ across the Hawaiian Islands, resulting in a NPP-to-GPP ratio ranging from 0.37 - 0.56.

>>>>

## Abstract
The State of Hawaiʻi passed legislation to be carbon neutral by 2045, a goal that will partly depend on carbon sequestration by terrestrial ecosystems. However, there is considerable uncertainty surrounding the future direction and magnitude of the land carbon sink in the Hawaiian Islands. We used the Land Use and Carbon Scenario Simulator (LUCAS), a spatially explicit stochastic simulation model that integrates landscape change and carbon gain-loss, to assess how projected future changes in climate and land use will influence ecosystem carbon balance in the Hawaiian Islands under all combinations of two radiative forcing scenarios (RCPs 4.5 and 8.5) and two land use scenarios (low and high) over a 90-year timespan from 2010-2100. Collectively, terrestrial ecosystems of the Hawaiian Islands acted as a net carbon sink under low radiative forcing (RCP 4.5) for the entire 90-year simulation period, with low land use change further enhancing carbon sink strength. In contrast, Hawaiian terrestrial ecosystems transitioned from a net sink to a net source of CO~2~ to the atmosphere under high radiative forcing (RCP 8.5), with high land use accelerating this transition and exacerbating net carbon loss. A sensitivity test of the CO~2~ fertilization effect on plant productivity revealed it to be a major source of uncertainty in projections of ecosystem carbon balance. Greater mechanistic understanding of plant productivity responses to rising atmospheric CO~2~ will be essential to realistically constrain simulation models used to evaluate the effectiveness of ecosystem-based climate mitigation strategies.

@@

The State of Hawaiʻi exemplifies the challenges associated with projecting the interactive effects of future climate and land use change on ecosystem carbon balance at a regional scale. Hawaiʻi was the first U.S. state to enact legislation committing to full carbon neutrality, requiring the state to account for and offset all of its greenhouse gas emissions by 2045 (State of Hawaiʻi Acts 15 and 16). This legislation emphasizes the mitigation potential of natural ecosystems as a key component to emissions reduction, necessitating baseline estimates and future projections of land carbon sink strength. However, Hawaiʻi's challenging terrain complicates these assessment efforts. The main Hawaiian Islands are a complex mosaic of natural and human-dominated landscapes overlain by steep climate gradients across relatively short distances, with mean annual temperature ranging from ~4-24° C [@giambelluca_evapotranspiration_2014] and mean annual rainfall ranging from ~200-10,200 mm [@giambelluca_online_2013]. Temperatures have risen rapidly in the Hawaiian Islands since the mid 1970s [@giambelluca_secular_2008; @mckenzie_temperature_2019] and a long-term drying trend has persisted since the early 1920s [@frazier_spatial_2017], resulting in reduced forest biomass and productivity [@barbosa_effects_2017]. These same drying and warming trends have increased the frequency and intensity of wildland fire [@trauernicht_contemporary_2015; @trauernicht_vegetationrainfall_2019] with predictable negative effects on ecosystem carbon balance [@selmants_baseline_2017]. Ecosystem carbon stocks across the main Hawaiian Islands have also been strongly influenced by the legacy of past land use change [@osher_effect_2003; @asner_high-resolution_2011]. Thousands of hectares were deforested beginning in the late 19th century to clear land for sugar plantations and cattle pasture [@cuddihy_alteration_1990]. Since the mid-20th century, much of this agricultural land has been steadily converted to urban areas, commercial forestry plantations, or simply abandoned and colonized by non-native grass and shrub species [@suryanata_diversified_2009; @perroy_evolving_2016]. Although these past trends surely inform the future impact of climate and land use change on ecosystem carbon balance, high spatial and temporal heterogeneity complicates realistic projection efforts. To date only one study has attempted to integrate land use, climate, and natural disturbances into future projections of Hawaiian ecosystem carbon balance, with projections limited to the mid-21st century under a single land use change scenario and a single moderate radiative forcing scenario [Special Report on Emission Scenarios [SRES] A1B, equivalent to Representative Concentration pathway [RCP] 6; @selmants_baseline_2017].

We used a stochastic, spatially explicit simulation model to estimate ecosystem carbon balance for Hawaiʻi's natural and agricultural lands on an annual basis for the period 2010–2100 under a range of assumptions about future climate, land use, land cover, disturbance, and global CO~2~ emissions [@daniel_state-and-transition_2016; @daniel_integrating_2018; @sleeter_effects_2019]. We developed four unique scenarios that explored different pathways, or future trajectories, of land use and climate change. These four scenarios represent all combinations of two land use change pathways (low and high) and two radiative forcing pathways (representative concentration pathway [RCP] 4.5 and RCP 8.5). In addition to these four scenarios, we conducted a separate series of simulations to examine how ecosystem carbon balance estimates vary in response to different levels of a CO~2~ fertilization effect (CFE) on net primary productivity [NPP; @sleeter_effects_2019]. Our goals were to estimate statewide changes in Hawaiian ecosystem carbon balance and their uncertainties under a range of plausible future scenarios, quantify the relative impact of major controlling processes such as land use change, disturbance, and climate change, and assess the sensitivity of modeled ecosystem carbon balance estimates to varying levels of a CFE on NPP.

@@

### *2.1 Study area*
The spatial extent of this study was the terrestrial portion of the seven main Hawaiian Islands (figure 1), a total land area of 16,554 km^2^. We subdivided this landscape into a grid of 264,870 simulation cells, each of which was 250 x 250 m in size. Each simulation cell was assigned to one of 210 possible state types based on the unique combination of seven islands, three moisture zones (dry, mesic, and wet; supplemental figure 1), and ten discrete land cover classes (figure 1). We also tracked the age and time since transition for each simulation cell as continuous state variables [@daniel_state-and-transition_2016].

```{r fig.align = 'center', out.width = "99%", fig.cap = "Land cover classification of the seven main Hawaiian Islands, adapted from Jacobi \\textit{et al.} (2017). Agriculture in this map combines herbaceous and woody crops, but these two crop types are treated as separate land cover classes in the simulation model. Water and Wetland land cover classes are not shown."}
knitr::include_graphics("./fig_images/fig1_HI_LC.png")
```  
### *2.2 States and transitions*
We developed two land use scenarios (low and high) with transition pathways modified from @daniel_state-and-transition_2016. Transition pathways were pre-defined to represent urbanization, agricultural contraction, agricultural expansion, harvesting of tree plantations, and wildfire. Agriculture, forest, grassland, shrubland, and tree plantation land cover classes each had multiple transition pathways, i.e., they could each be converted into a variety of other land cover classes. The barren land cover class could only transition to developed, so its sole transition pathway was urbanization. Although most land cover classes had an urbanization transition pathway, there was no transition pathway out of an urbanized (developed) state. Water and wetland land cover classes remained static throughout the simulation period.

Transition targets set the area to be transitioned over time. These targets were based on either recent historical trends of land use change in the Hawaiian Islands from 1992-2011 [@noaa_coastal_2020] or on population projections for the State of Hawaiʻi [@kim_population_2018]. For the high land use scenario, rates of agricultural contraction, agricultural expansion, and urbanization for each timestep and Monte Carlo realization were sampled from uniform distributions bounded by the median and maximum recent historical rates for each island. For the low land use scenario, rates of agricultural contraction and expansion were sampled from uniform distributions bounded by zero and the minimum recent historical rates for each island, which shifts the balance toward agricultural contraction and leads to a reduction in land area under active cultivation over time. Urbanization rates in the low land use scenario were based on island-level population estimates and projections at five year intervals from 2010-2045 [@kim_population_2018]. We converted population projections into urbanization transition targets following @sleeter_future_2017 by calculating population density for each island and then projecting future developed area based on the five-year incremental change in island population. The spatial extent of agricultural contraction, agricultural expansion, and urbanization was constrained in both land use scenarios based on existing zoning maps [@daniel_state-and-transition_2016]. Transition targets for tree plantation harvest were set at ~75% of recent historical rates in the high land use scenario and ~40% of recent historical rates in the low land use scenario [@daniel_state-and-transition_2016]. Tree plantation forestry in the State of Hawaiʻi is primarily short-rotation (5-7 years) *Eucalyptus* spp., harvested at a rate of approximately 2 km^2^ y^-1^ over the past decade. In both the high and low land use scenarios, approximately 60% of tree plantation harvests were replacement harvests resulting in conversion to grassland or agriculture. The remaining 40% were rotation harvests replanted to *Eucalyptus* spp. There was no transition pathway from any land cover class into tree plantation in either land use scenario, which is consistent with recent historical trends of stable or declining tree plantation land area since the year 2000. For the contemporary period (2010-2020), transition targets in both the high and low land use scenarios were set at low land use scenario rates and diverged after 2020.

@@

Initial carbon stocks and baseline carbon flows were derived from the Integrated Biosphere Simulator [IBIS; @foley_integrated_1996; @liu_critical_2020], a process-based dynamic global vegetation model. We initiated IBIS with minimal vegetation and simulated forward for 110 years using 30-year climate normals for the Hawaiian Islands [supplemental Figure 3; @giambelluca_online_2013; @giambelluca_evapotranspiration_2014]. We calibrated IBIS carbon stocks with statewide gridded datasets of soil organic carbon [@ssurgo_2016] and total forest live biomass. The total forest live biomass data was derived from @asner_rapid_2016 estimates of aboveground forest live biomass by applying a power function to calculate the belowground portion [@mokany_critical_2006] as described in @selmants_baseline_2017. We calibrated gross photosynthesis in IBIS using a Hawaiʻi-specific gridded dataset of gross primary production (GPP) derived from MODIS satellite imagery [@kimball_evaluating_2017]. Mean annual GPP from @kimball_evaluating_2017 ranged from 0.123 - 3.88 kg C m^-2^ y^-1^ across the Hawaiian Islands, resulting in a NPP-to-GPP ratio ranging from 0.37 - 0.56.
